---
title: "Is Your Electric Vehicle Overpriced?"
author: "Natalie Stier"
date: "2024-01-26"
output:
  slidy_presentation:
    css: bootstrap.min.css
runtime: shiny
resource_files:
- bootstrap.min.css
- ecars.csv
- model.rds
- global.R
- most_makes.csv
- predicted_price.csv
- bootstrap.min.css
- ecars.csv
- global.R
- model.rds
- most_makes.csv
- predicted_price.csv
- bootstrap.min.css
- ecars.csv
- global.R
- model.rds
- most_makes.csv
- predicted_price.csv
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(tidyr)
library(car)
library(MASS)
library(repr)
library(pals)
library(shinythemes)
library(plotly)
library(bslib)
library(ggpubr)
library(revealjs)
library(shiny)
library(extrafont)


ecars = read.csv('ecars.csv')
ecars_vars = ecars[,4:10]
ecars_make  = ecars %>% group_by(Make) %>%
  filter(n()>=10)
most_makes = read.csv('most_makes.csv')
predicted_price = read.csv('predicted_price.csv')

makes = most_makes$Make 

ecars$Make2 = ifelse(ecars$Make %in% makes, ecars$Make, "zOther")
make_other = unique(ecars$Make2)
make_other = sort(make_other)
make_other[15] = 'Other'

price_model_initial = lm(Price ~ Efficiency + Fast_charge + 
                        Range + Top_speed + 
                        Acceleration, data = ecars)
bc = boxCox(price_model_initial)
lambda = bc$x[which(bc$y == max(bc$y))]

model <- readRDS("model.rds")
             
make_colors = c('#e6194b', '#f58231',  '#ffe119', 
                '#bcf60c','#3cb44b', '#008080',
                '#aaffc3', '#4363d8', '#89A8BF',
                '#46f0f0', '#911eb4', '#e6beff',
                '#f032e6', '#fabebe')
              

make_colors2 = c('#e6194b', '#f58231',  '#ffe119', 
                '#bcf60c','#3cb44b', '#008080',
                '#aaffc3', '#4363d8', '#89A8BF',
                '#46f0f0', '#911eb4', '#e6beff',
                '#f032e6', '#fabebe', 'black')

make_other = unique(ecars$Make2)
make_other = sort(make_other)
make_other[15] = 'Other'
make_other

```

## Research Question and Background Information:

-   Research Question:

Based on an electric model's features what should its price be?

-   Why Electric Cars?

I wanted to work with a data set lending itself to a multiple linear regression.

According to [Autoweek](https://www.autoweek.com/news/a44900159/are-electric-cars-worth-it/) electric cars are typically \$6000 to \$7000 more than their gas counterparts:

> Some electric cars are the same or less money than their gas counterparts, but others are much more expensive.\
> It all depends on the car, the price of electricity, the rate of depreciation, if you're eligible for any incentives, and much more.

**The more you know about your investment the better.**

## The Data Set

This collection of data on electric vehicles came from [Kaggle](https://www.kaggle.com/datasets/fatihilhan/electric-vehicle-specifications-and-prices/). The original set included 360 observations and 9 variables.

**Variables**

-   Battery: The capacity of the vehicle's battery in kilowatt-hours (kWh).
-   Car_name: The model name of the electric vehicle.
-   Car_name_link: A direct link to the corresponding page on EV Database for more in-depth information.
-   Efficiency: The energy efficiency rating of the vehicle in watt-hours per kilometer (Wh/km).
-   Fast_charge: The fast-charging capability of the vehicle in range per hour of fast charging..
-   Price.DE.: The price of the electric vehicle in Germany.
-   Range: The driving range of the vehicle on a single charge in kilometers.
-   Top_speed: The maximum speed the vehicle can achieve in kilometers per hour.
-   Acceleration..0.100.: The acceleration time from 0 to 100 kilometers per hour.

## Data Processing and Variables

<div style= "float:right;position: relative; top: 0px;">
```{r, echo = FALSE, fig.dim = c(8, 6)}
 
output$features <- renderPlot({
    plot(ecars[,4:10], 
         main = 'Comparison of all Continuous Variables', 
         cex =1.5)  
  })

plotOutput("features", height = '650px', width = '800px')
```
</div>

1.  I created a Make variable by extracting the first word from the Car_name variable.
2.  I also renamed several columns to make them more intuitive. For example, I changed acceleration..0.100. to Acceleration.
3.  I split the dataframe into two. A dataframe containing cars with prices called ecars (309 objects), and one containing cars missing prices called ecars_missing_price (51 objects).
4.  I removed the two cars that did not have fast charge capability (the Renault Twingo Electric and the e.Go e.wave X) because this was an important feature in the linear regression and was impacting their price.

307 electric cars with 45 unique car makes were included in the ecars data used to create the linear model. Additionally 14 makes that have 10 or more car models are highlighted throughout the project.

## Quantitative Variables

```{r, echo = FALSE}
inputPanel(
  varSelectInput("var", 
                 tags$span(style="color: #343a40; font-size: 24px;","Feature"), 
                 ecars[,4:10], 
                 selected = "Price")
  )

output$p1 <- renderPlotly({
  plot_ly(
      data = ecars,
      x = ~get(input$var),
      type = "box",
      text = ~Car_name,
      name = ' ',
      tooltip = c("x", "text")) %>% 
      layout(title = "Box Plot of Selected Feature",
             yaxis = list(title = ''),
             xaxis = list(title = ''))
   
  })
column(4,
plotlyOutput('p1', width = '500px', height = '500px')
)

  output$p2 <- renderPlotly({
     plot_ly(
      data = ecars_make,
      x = ~get(input$var),
      y = ~Make,
      type = "box",
      color = ~Make,
      colors = make_colors,
      text = ~Car_name,
      tooltip = c("x", "text"),
      showlegend = FALSE
              )%>% 
      layout(title = "Selected Feature (Makes with 10 + Models)",
             yaxis = list(title = ''),
             xaxis = list(title = ''))
    
  })

column(4,
plotlyOutput('p2', width = '600px', height = '500px')
)
  
 output$histogram <- renderPlot({
    ggplot(ecars, aes(x=!!input$var)) + 
      geom_histogram(color="black", fill="white", bins = 15)+
      ggtitle('Distribution of Selected Feature')+ 
      theme(plot.title = element_text(size=18, family= "Verdana"),
            axis.title = element_text(size = 14))  
            
  })  
 
column(4, 
       plotOutput('histogram', width = '500px', height = '500px')
)

```

## Quantitative Variable Relationships

```{r, echo = FALSE}
column(2, 
      tags$label(h3('Select Variables')),
      varSelectInput("xvar", 
                     "X variable", 
                     ecars_vars, 
                     selected = "Battery"),
      varSelectInput("yvar", 
                     "Y variable", 
                     ecars_vars, 
                     selected = "Price")
        )
column(8, 
       plotOutput("scatter", height = '600px')
        )
column(2, 
            checkboxGroupInput("make", "Filter by Make",
                                choices = unique(ecars_make$Make), 
                                selected = unique(ecars_make$Make)),
            hr(), # Add a horizontal rule
            checkboxInput("by_make", "Show Make", TRUE),
            checkboxInput("smooth", "Add Smoother")
            )

subsetted <- reactive({
    req(input$var)
    ecars |> filter(Make %in% input$make)
  })
  
output$scatter <- renderPlot({
    p <- ggplot(subsetted(), aes(!!input$xvar, !!input$yvar)) + 
      scale_color_manual(values = make_colors) +
      list(theme(legend.position = "bottom"),
      if (input$by_make) aes(color = Make),
      geom_point(), 
      if (input$smooth) geom_smooth()
    )
    
    p
    
  }, res = 100)

```

## Top Speed vs. Price

Top speed is the single best predictor of price among the quantitative variables with an R-squared value of 0.5771 and a p-value of 2.2e-16 for both coefficients. 

```{r, echo = FALSE}

 output$makescatter <- renderPlotly({

test = ggplot(ecars, aes(x = Top_speed, y = Price, text = Car_name)) +
  geom_point(aes(col = Make2)) + 
  scale_color_manual(name = "Make", values = make_colors2, labels = make_other)+
  xlab("Top Speed (km/hr)") + 
  ylab("Price in Germany (euros) ") +
  ggtitle('Top Speed vs. Price (Makes with 10+ Models Highlighted)') +
  theme(legend.position = "none")
ggplotly(test, tooltip = c("x", 'y', "text")) 
 })
 
output$barchart <- renderPlot({
ecars %>% group_by(Make) %>%
    filter(n() >= 10) %>%
    summarise(feature = (mean(Price))) %>%
    ggplot(., aes(x = Make, y = feature, fill = Make, label = Make)) +
    geom_bar(stat = 'identity') +
    scale_fill_manual(values = make_colors) +
    ylab('Mean')+
    theme(axis.title.x=element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank()) +
    geom_text(angle = 90, position = position_stack(vjust = 0.5)) + 
    theme(legend.position = "none", plot.title = element_text(size=18))+
    ggtitle('Mean Price for each Make with 10+ Models')
})

column(8, plotlyOutput('makescatter', width = '900px', height = '600px'))
column(4, plotOutput('barchart', height = '600px'))

```

```{r}
#lm_Price_TS = lm(Price ~ Top_speed, ecars)
#summary(lm_Price_TS)
```

## Multiple Linear Regression

<div style= "float:right;position: relative; top: 0px;">
```{r, echo = FALSE}
summary(model)
```
</div>

1.  I started by making a linear model including all the continuous variables.
2.  Using that model and an empty model I did an AIC which eliminated Battery. I expected either Battery or Range to be eliminated because of their co-linearity.
3.  Price is skewed right so I did a BoxCox transformation and then redid the model after transforming the price.

## Linear Model

The predicted price of each car is plotted against its actual price to easily visualize which vehicles cost more vs. less than predicted.

The blue line indicates the predicted price for all vehicles in the data set. The black dashed lines indicate the confidence interval and the the red dashed lines indicate the prediction interval.

```{r, echo = FALSE}
column(2,
      sliderInput('xzoom', 'x zoom', 0, 300, 250, ticks = TRUE),
      sliderInput('yzoom', 'y zoom', 0, 300, 250, ticks = TRUE))
column(10,
      plotlyOutput('MLR', width = '1000px', height = '550px'))

output$MLR <- renderPlotly({
    mlr = ggplot(NULL, aes(Predicted_price, Price)) +
      geom_smooth(data = predicted_price, aes(x = Predicted, y = Predicted), 
                  col = 'blue', size = .8, alpha = .5) +
      geom_smooth(data = predicted_price, aes(x = Predicted, y = Predict_lwr), 
                  col = 'red', linetype = 'dashed', size = .8, alpha = .8) +
      geom_smooth(data = predicted_price, aes(x = Predicted, y = Predict_upr), 
                  col = 'red', linetype = 'dashed',size = .8, alpha = .8) +
      geom_smooth(data = predicted_price, aes(x = Predicted, y = Confidence_lwr), 
                  col = 'black', linetype = 'dashed', size = .8, alpha = .8) +
      geom_smooth(data = predicted_price, aes(x = Predicted, y = Confidence_upr), 
                  col = 'black', linetype = 'dashed', size = .8, alpha = .8) +
      geom_point(data = predicted_price, aes(x = Predicted, y = Price, text = Name), alpha = .5) +
      theme(legend.position = "right", legend.text = element_text(size = 8))+
      ylim(25, input$yzoom) + xlim(25, input$xzoom) +
      labs(title = "Predicted Price vs. Price for all EV Models",
           caption = "Data source: ToothGrowth",
           x = "Predicted Price (euros in thousands)", y = "German Price (euros in thousands)",
           tag = "A")
    
    ggplotly(mlr, tooltip = c("x", 'y', "text")) 
  })
 
```

## Price Prediction using the Linear Model

```{r, echo = FALSE}

  datasetInput <- reactive({  
    
    df <- data.frame(Efficiency = input$eff,
                      Fast_charge = input$fc, 
                      Range = input$range, 
                      Top_speed = input$ts, 
                      Acceleration = input$acc,
                      stringsAsFactors = FALSE)
   
    input <- df
    
    Output <- (((predict(model, df)) *lambda) + 1)^(1/lambda)
    
    print(Output)
    
  })
  
  # Status/Output Text Box
  output$contents <- renderPrint({
    if (input$submitbutton>0) { 
      isolate("Calculation complete.") 
    } else {
      return("Server is ready for calculation.")
    }
  })
  
  # Prediction results table
  output$tabledata <- renderTable({
    if (input$submitbutton>0) { 
      isolate(datasetInput()) 
    } 
  })
output$pm <- renderPlot({
    df <- data.frame(Efficiency = input$eff,
                     Fast_charge = input$fc, 
                     Range = input$range, 
                     Top_speed = input$ts, 
                     Acceleration = input$acc,
                     stringsAsFactors = FALSE)
    ap = input$ap 
    ymax = input$ylab
    xmax = input$xlab
    
    input <- df
    
    pp = as.numeric(((predict(model, df)) *lambda) + 1)^(1/lambda)   
    
    df2 = data.frame(x1 = pp,
                     y1 = ap)
    
    ggplot(NULL, aes(Predicted_price, Price)) +
      geom_smooth(data = predicted_price, aes(x = Predicted, y = Predicted), col = 'blue', size = .5, alpha = .8) +
      geom_smooth(data = predicted_price, aes(x = Predicted, y = Predict_lwr), col = 'red', linetype = 'dashed', alpha = .8) +
      geom_smooth(data = predicted_price, aes(x = Predicted, y = Predict_upr), col = 'red', linetype = 'dashed', alpha = .8) +
      geom_smooth(data = predicted_price, aes(x = Predicted, y = Confidence_lwr), col = 'black', linetype = 'dashed', alpha = .8) +
      geom_smooth(data = predicted_price, aes(x = Predicted, y = Confidence_upr), col = 'black', linetype = 'dashed', alpha = .8) +
      geom_point(data = predicted_price, aes(x = Predicted, y = Price), alpha = .7) +
      geom_point(data = most_makes, aes(x = mean_predicted, y = mean_price), size = 4) +
      geom_point(data = most_makes, aes(x = mean_predicted, y = mean_price, col = Make), size = 3) +
      geom_point(data = df2, aes(x = x1/1000, y = y1/1000), size = 7, pch= 18, col = 'black') +
      geom_point(data = df2, aes(x = x1/1000, y = y1/1000), size = 6, pch= 18, col = 'yellow') +
      geom_point(data = df2, aes(x = x1/1000, y = y1/1000), size = 3, pch= 18, col = 'black') +
      theme(legend.position = "bottom", legend.text = element_text(size = 10))+
      labs(col = 'Car Make') +
      ylim(25, ymax) + xlim(25, xmax) +
      scale_color_manual(values = make_colors) +
      xlab("Predicted Price (euros in thousands)") + ylab("Price (euros in thousands)") +
      ggtitle('Predicted Price vs. Price', subtitle = 'Means for Makes with 10+ Models Included')+ 
      theme(plot.title = element_text(size = 18), 
            plot.subtitle = element_text(size = 14),
            axis.title = element_text(size = 14))
    

    
  })

column(2, 
 numericInput("eff", 
                          label = "Efficiency (Wh/km)", 
                          value = 200),
            numericInput("fc", 
                          label = "Fast Charge (km/hr)", 
                          value = 500),
            numericInput("range", 
                          label = "Range (km)", 
                          value = 350),
            numericInput('ts', 
                          label = "Top Speed (km/hr)", 
                          value = 150),
            numericInput('acc', 
                         label = "Acceleration (s to 100 km/hr)", 
                         value = 6),
            numericInput('ap', 
                         label = "Price (euros)", 
                         value = 50000)
            
        )

column(2, 
       'Hit submit for',
       'the predicted price',
       actionButton("submitbutton",
                    "Submit", 
                     class = "btn btn-primary \n"),
       tableOutput('tabledata'),
       sliderInput('xlab', 'x zoom', 0, 500, 150, ticks = TRUE),
       sliderInput('ylab', 'y zoom', 0, 500, 150, ticks = TRUE))
column(8,
       plotOutput('pm', height = '650px'))
```

## Conclusions and Future Work 

<div style= "float:right;position: relative; top: 0px;">
```{r, echo = FALSE}
MM = most_makes
MM = mutate(MM, Difference = (mean_price - mean_predicted))
MM = rename(MM, Mean.Price = mean_price, Mean.Predicted = mean_predicted)
MM
```
</div>

Conclusions:  

- Mercedes, BMW, and, in particular, Porsche models, are priced higher than predicted.

- MG models are a good value based on their price and features.  

- Audi, Volkswagen, Citroen, Peugeot, and Volvo models are priced very close to tge predicted price. 

- Tesla models are priced slightly below their predicted value. 

- Both the Lotus Eletre models are priced about 85,000 euros less than predicted. 

- The Lucid Air Dream Editions P & R are priced almost 60,000 euros more than predicted

Future Work:

- It would be interesting to include other features in the linear model. For example, the Make of the car likely makes a significant difference in the price.

