---
title: "Is Your Electric Vehicle Overpriced?"
author: "Natalie Stier"
output: 
  slidy_presentation:
    css: bootstrap.min.css
runtime: shiny
date: "2024-01-08"
---

```{r setup, include=FALSE}
library(bslib)
library(dplyr)
library(ggplot2)
library(tidyr)
library(car)
library(MASS)
library(repr)
library(pals)
library(ggpubr)
library(plotly)
library(revealjs)
library(shiny)
knitr::opts_chunk$set(echo = FALSE)

ecars <- read.csv("ecars.csv")

top_10 = ecars %>% group_by(Make) %>%
  filter(n() >= 10)


makes = top_10$Make 

ecars$Make2 = ifelse(ecars$Make %in% makes, ecars$Make, "zOther")
make_other = unique(ecars$Make2)
make_other = sort(make_other)
make_other[15] = 'Other'
make_colors = c('#e6194b', '#f58231',  '#ffe119', 
                '#bcf60c','#3cb44b', '#008080',
                '#aaffc3', '#4363d8', '#000075',
                '#46f0f0', '#911eb4', '#e6beff',
                '#f032e6', '#fabebe')
make_colors2 = c('#e6194b', '#f58231',  '#ffe119', 
                '#bcf60c','#3cb44b', '#008080',
                '#aaffc3', '#4363d8', '#000075',
                '#46f0f0', '#911eb4', '#e6beff',
                '#f032e6', '#fabebe', 'black')


```

## Background Information and Research Question:

I wanted to focus on linear regression. 
Electric cars are typically $6000 to $7000 more than their gas counterparts. So it is already more of an investment. 


## The data set 

This collection of data on electric vehicles came from Kaggle. The original set included 360 observations and 9 variables.

https://www.kaggle.com/datasets/fatihilhan/electric-vehicle-specifications-and-prices/ 

Data Fields:
Battery: The capacity of the vehicle's battery in kilowatt-hours (kWh).
Car_name: The model name of the electric vehicle.
Car_name_link: A direct link to the corresponding page on EV Database for more in-depth information.
Efficiency: The energy efficiency rating of the vehicle in watt-hours per kilometer (Wh/km).
Fast_charge: The fast-charging capability of the vehicle in minutes for a certain charging percentage.
Price.DE.: The price of the electric vehicle in Germany.
Range: The driving range of the vehicle on a single charge in kilometers.
Top_speed: The maximum speed the vehicle can achieve in kilometers per hour.
Acceleration..0.100.: The acceleration time from 0 to 100 kilometers per hour.

## Qualitative Variable included in the data set. 

```{r}
plot(ecars[,4:10])
```

## Slide with Plot

```{r}
make_colors = c('#e6194b', '#f58231',  '#ffe119', 
                '#bcf60c','#3cb44b', '#008080',
                '#aaffc3', '#4363d8', '#000075',
                '#46f0f0', '#911eb4', '#e6beff',
                '#f032e6', '#fabebe')


test = ggplot(ecars, aes(x = Battery, y = Price, text = Car_name)) +
  geom_point(aes(col = Make2)) + 
  scale_color_manual(name = "Make", values = make_colors2, labels = make_other)+
  xlab("Battery Capacity (kWh)") + 
  ylab("Price in Germany (euros) ") +
  ggtitle('Electric Vehicle Battery vs. Price (Makes with 10+ Models Highlighted)') +
  theme(legend.position = "none")


ggplotly(test, tooltip = c("x", 'y', "text")) 
```

## Page plot
<div style= "float:right;position: relative; top: -80px;">


```{r}
a = ggplot(ecars, aes(x = Battery, y = Price, text = Car_name)) +
  geom_point(aes(col = Make2)) + 
  scale_color_manual(name = "Make", values = make_colors2, labels = make_other)+
  xlab("Battery Capacity (kWh)") + 
  ylab("Price in Germany (euros) ") +
  ggtitle('Electric Vehicle Battery vs. Price (Makes with 10+ Models Highlighted)') +
  theme(legend.position = "none")

b =  ggplot(ecars, aes(x = Acceleration, y = Price, text = Car_name)) +
  geom_point(aes(col = Make2)) + 
  scale_color_manual(name = "Make", values = make_colors2, labels = make_other)+
  xlab("Battery Capacity (kWh)") + 
  ylab("Price in Germany (euros) ") +
  ggtitle('Electric Vehicle Battery vs. Price (Makes with 10+ Models Highlighted)') +
  theme(legend.position = "none")


ggarrange(a, b, 
          labels = c("A", "B"),
          nrow = 2,
          widths = c(1, 1))
```
</div>
THis is the text
It is showing that 
the is a linear model 
hello 
why are we doing this this is thisfijdfsdf
fjsdlkfhelel adfalehffe hhfjdhf 
HIIIIII


## Shiny

```{r}
inputPanel(
  varSelectInput("var", 
                 "Feature", 
                 ecars[,4:10], 
                 selected = "Price")
  )


output$p1 <- renderPlot({
  ggplot(top_10, aes(!!input$var, y = factor(0))) +
      geom_boxplot()+
      theme(axis.title.y=element_blank(),
            axis.text.y=element_blank(),
            axis.ticks.y=element_blank()) +
      ggtitle('Box Plot of Selected Feature') 
})

output$p2 <- renderPlot({
  ggplot(top_10, aes(!!input$var, Make, fill = Make)) +
      scale_fill_manual(values = make_colors) +
      geom_boxplot(outlier.colour="black", 
                   outlier.shape=16, 
                   outlier.size=2, 
                   notch=FALSE) +
      theme(legend.position = "none") +
      ggtitle('Box Plots of Selected Feature by Makes with 10 or More Models')
})

column(6,
plotOutput('p1', width = '600px')
)

column(6,
plotOutput('p2', width = '600px')
)
```


## another
```{r}

 
column(2, 
    varSelectInput("var", 
                   "Feature", 
                   ecars[,4:10], 
                   selected = "Price")
    
   )
column(6,
          plotOutput("box")
   )
column(6,
          plotOutput("box2")
)
 
```


